KidsAIR: Main analysis initial steps

### 2020-10-26 main_analysis_initial.Rmd
analysis updates

lrti_final_data <- read_rds(paste0(file_path, "Output/lrti_final_data.rds")) %>% 
  arrange(area, home, child_id_num, winter_id) %>% 
  mutate(cohort = round_date(csq_date, unit = "year")) %>% 
  separate(cohort, c("cohort", "trash"), sep = "-") %>% 
  group_by(area, child_id_num, winter_id) %>% 
  mutate(lrti_events = n()) %>% 
  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
  group_by(area, child_id_num) %>% 
  mutate(lrti_events = as.numeric(lrti_events),
         lrti_events = if_else(is.na(person_time_no_lrti_total), 0, lrti_events),
         lrti_events_overall = sum(lrti_events, na.rm = TRUE),
         person_time_no_lrti_total = as.numeric(person_time_no_lrti_total),
         person_time_no_lrti_total = if_else(is.na(person_time_no_lrti_total), person_time_possible,
                                             person_time_no_lrti_total),
         person_time_at_risk_total = sum(person_time_no_lrti_total, na.rm = TRUE),
         person_time_at_risk_total = as.numeric(person_time_at_risk_total)/7,
         lrti_incidence = lrti_events_overall/person_time_at_risk_total,
         lrti_incidence = if_else(is.na(lrti_incidence), 0, lrti_incidence)) %>% 
  arrange(area, home, child_id_num, winter_id) %>% 
  left_join(kids_amb_pm_data, by = c("area", "home")) %>% 
  mutate(pm_filter = if_else(sample_date >= person_time_start & sample_date <= person_time_end, 1, 0)) %>% 
  filter(pm_filter == 1) %>% 
  group_by(area, child_id_num) %>% 
  mutate(amb_pm_lrti = mean(amb_pm_24hr, na.rm = TRUE)) %>% 
  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
  ungroup() %>% 
  select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence,
         amb_pm_lrti, person_time_start, person_time_end) %>% 
  left_join(kids_amb_temp_data, by = c("area", "home")) %>% 
  mutate(temp_filter = if_else(sample_date >= person_time_start & sample_date <= person_time_end, 1, 0)) %>% 
  filter(temp_filter == 1 | child_id_num == 94) %>% 
  group_by(area, child_id_num) %>% 
  mutate(amb_temp_lrti = mean(mean_temp, na.rm = TRUE),
         amb_temp_lrti = if_else(child_id_num == 94, NaN, amb_temp_lrti),
         amb_dewp_lrti = mean(mean_dewp, na.rm = TRUE),
         amb_dewp_lrti = if_else(child_id_num == 94, NaN, amb_dewp_lrti),
         amb_pressure_lrti = mean(mean_pressure, na.rm = TRUE),
         amb_pressure_lrti = if_else(child_id_num == 94, NaN, amb_pressure_lrti),
         amb_wind_lrti = mean(mean_wind_knots, na.rm = TRUE),
         amb_wind_lrti = if_else(child_id_num == 94, NaN, amb_wind_lrti)) %>% 
  distinct(area, child_id_num, .keep_all = TRUE) %>% 
  ungroup() %>% 
  select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence,
         amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti)

# Merge dataset with blinded treatments and final LRTI data, select analysis variables
kids_analysis_data_1obs_per_child <- health_exposure_data_sampling_day %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  mutate(child_age = fct_collapse(age_child_nov,
                                  "< 1" = "Less than 1",
                                  "1+" = c(1, 2, 3, 4))) %>% 
  left_join(treatments_masked, by = "home") %>% 
  arrange(area, home, child_id_num, winter_id, age_child_nov) %>% 
  right_join(lrti_final_data, by = c("area", "home", "child_id_num")) %>% 
  distinct(area, child_id_num, .keep_all = TRUE) %>% 
  select(area, home, cohort, child_id_num, child_age, lrti_events_overall, 
         person_time_at_risk_total, lrti_incidence, newtx,
         amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti,
         gender_parent, hispanic_parent, race_parent, education_3level, income_3level,
         home_floors_2level, home_year_built_2level, home_pets_2level, home_pets,
         stove_age_3level, chimney_clean_3level, wood_collect_method_2level,
         wood_collect_2level, stove_grade_3level, burn_level_3level, ruca_code,
         ruca_code_2, ruca_code_3, residents_smoke, residents_smoke_inside,
         residents_under_five, home_bedrooms, pm_mean_sampling_period, pm_sample_interval,
         moisture_closest, temp_indoor_max, rh_indoor_max, sums_events_sampling_period_2.5,
         sums_events_sampling_period_5, sums_events_sampling_period_7.5) %>% 
  rename(treatment = newtx,
         lrti_events = lrti_events_overall,
         person_time_at_risk = person_time_at_risk_total) %>% 
  mutate(log_person_time_at_risk = log(person_time_at_risk)) %>% 
  mutate(lrti_events_di = if_else(lrti_events == 0, 0, 1)) %>% 
  mutate(treatment = factor(treatment,
                            levels = c("Blackfoot", "Dearborn", "Smith")))

@@

analysis_data <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_1obs_per_child.rds"))

@@

# amb data summary
# amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti
ambient_summary <- analysis_data %>% 
  filter(area == "NN") %>% 
  mutate(new_var = amb_pm_lrti) %>% 
  group_by(treatment) %>% 
  filter(!is.na(new_var)) %>% 
  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% 
  summarize(mean_amb = mean(new_var, na.rm = TRUE),
            sd_amb = sd(new_var, na.rm = TRUE),
            n_amb = n(),
            min_amb = min(new_var, na.rm = TRUE),
            median_amb = median(new_var, na.rm = TRUE),
            max_amb = max(new_var, na.rm = TRUE))
ambient_summary

# Histogram for distribution of data
events_plot <- analysis_data %>% 
  ggplot() +
    geom_histogram(aes(lrti_events)) +
    theme_light() +
    facet_wrap(~treatment)
events_plot

events_di_plot <- analysis_data %>% 
  ggplot() +
    geom_boxplot(aes(lrti_events_di, person_time_at_risk)) +
    facet_wrap(~treatment)
events_di_plot
```

```{r}
# Check for overdispersion in data using likelihood ratio test
library(pscl)
library(MASS)

nb_results <- glm.nb(lrti_events ~ treatment + offset(log_person_time_at_risk), 
                     data = analysis_data)

@@

odTest(nb_results)

# According to these results, data is overdispersed
# Variance is much higher than the mean for LRTI incidence
# Assumption of Poisson is variance = mean, so recommend using negative binomial model

summary(nb_results)

@@

# Other diagnostic plots and tests
library(vcd)
fit <- goodfit(analysis_data$lrti_events)
summary(fit) 
rootogram(fit)
Ord_plot(analysis_data$lrti_events)

library(car)
influencePlot(nb_results)

library(pscl)

@@

influence(nb_results)$coef
```

@@

plot(negbin_results)
qqnorm(residuals(negbin_results))
qqline(residuals(negbin_results))

@@

>>>>

lrti_final_data <- read_rds(paste0(file_path, "Output/lrti_final_data.rds")) %>% 
  arrange(area, home, child_id_num, winter_id) %>% 
  mutate(cohort = round_date(csq_date, unit = "year")) %>% 
  separate(cohort, c("cohort", "trash"), sep = "-") %>% 
  group_by(area, child_id_num, winter_id) %>% 
  mutate(lrti_events = n()) %>% 
  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
  group_by(area, child_id_num) %>% 
  mutate(lrti_events = as.numeric(lrti_events),
         lrti_events = if_else(is.na(person_time_no_lrti_total), 0, lrti_events),
         lrti_events_overall = sum(lrti_events, na.rm = TRUE),
         person_time_no_lrti_total = as.numeric(person_time_no_lrti_total),
         person_time_no_lrti_total = if_else(is.na(person_time_no_lrti_total), person_time_possible,
                                             person_time_no_lrti_total),
         person_time_at_risk_total = sum(person_time_no_lrti_total, na.rm = TRUE),
         person_time_at_risk_total = as.numeric(person_time_at_risk_total)/7,
         lrti_incidence = lrti_events_overall/person_time_at_risk_total,
         lrti_incidence = if_else(is.na(lrti_incidence), 0, lrti_incidence)) %>% 
  arrange(area, home, child_id_num, winter_id) %>% 
  left_join(kids_amb_pm_data, by = c("area", "home")) %>% 
  mutate(pm_filter = if_else(sample_date >= person_time_start & sample_date <= person_time_end, 1, 0)) %>% 
  filter(pm_filter == 1) %>% 
  group_by(area, child_id_num) %>% 
  mutate(amb_pm_lrti = mean(amb_pm_24hr, na.rm = TRUE)) %>% 
  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
  ungroup() %>% 
  dplyr::select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence,
         amb_pm_lrti, person_time_start, person_time_end) %>% 
  left_join(kids_amb_temp_data, by = c("area", "home")) %>% 
  mutate(temp_filter = if_else(sample_date >= person_time_start & sample_date <= person_time_end, 1, 0)) %>% 
  filter(temp_filter == 1 | child_id_num == 94) %>% 
  group_by(area, child_id_num) %>% 
  mutate(amb_temp_lrti = mean(mean_temp, na.rm = TRUE),
         amb_temp_lrti = if_else(child_id_num == 94, NaN, amb_temp_lrti),
         amb_dewp_lrti = mean(mean_dewp, na.rm = TRUE),
         amb_dewp_lrti = if_else(child_id_num == 94, NaN, amb_dewp_lrti),
         amb_pressure_lrti = mean(mean_pressure, na.rm = TRUE),
         amb_pressure_lrti = if_else(child_id_num == 94, NaN, amb_pressure_lrti),
         amb_wind_lrti = mean(mean_wind_knots, na.rm = TRUE),
         amb_wind_lrti = if_else(child_id_num == 94, NaN, amb_wind_lrti)) %>% 
  distinct(area, child_id_num, .keep_all = TRUE) %>% 
  ungroup() %>% 
  dplyr::select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence,
         amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti)

# Merge dataset with blinded treatments and final LRTI data, select analysis variables
kids_analysis_data_1obs_per_child <- health_exposure_data_sampling_day %>% 
  #select(-treatment_assigned, -treatment_actual) %>% 
  mutate(child_age = fct_collapse(age_child_nov,
                                  "< 1" = "Less than 1",
                                  "1+" = c(1, 2, 3, 4))) %>% 
  left_join(treatments_masked, by = "home") %>% 
  arrange(area, home, child_id_num, winter_id, age_child_nov) %>% 
  right_join(lrti_final_data, by = c("area", "home", "child_id_num")) %>% 
  distinct(area, child_id_num, .keep_all = TRUE) %>% 
  dplyr::select(area, home, cohort, child_id_num, child_age, lrti_events_overall, 
         person_time_at_risk_total, lrti_incidence, newtx, treatment_assigned,
         amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti,
         gender_parent, hispanic_parent, race_parent, education_3level, income_3level,
         home_floors_2level, home_year_built_2level, home_pets_2level, home_pets,
         stove_age_3level, chimney_clean_3level, wood_collect_method_2level,
         wood_collect_2level, stove_grade_3level, burn_level_3level, ruca_code,
         ruca_code_2, ruca_code_3, residents_smoke, residents_smoke_inside,
         residents_under_five, home_bedrooms, pm_mean_sampling_period, pm_sample_interval,
         moisture_closest, temp_indoor_max, rh_indoor_max, sums_events_sampling_period_2.5,
         sums_events_sampling_period_5, sums_events_sampling_period_7.5) %>% 
  rename(treatment = newtx,
         lrti_events = lrti_events_overall,
         person_time_at_risk = person_time_at_risk_total) %>% 
  mutate(log_person_time_at_risk = log(person_time_at_risk)) %>% 
  mutate(lrti_events_di = if_else(lrti_events == 0, 0, 1)) %>% 
  mutate(treatment = factor(treatment,
                            levels = c("Blackfoot", "Dearborn", "Smith")),
         treatment_assigned = factor(treatment_assigned,
                                     levels = c("Placebo", "Filter", "Education")))

@@

analysis_data <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_1obs_per_child.rds")) %>% 
  rownames_to_column() %>% 
  mutate(filter_outliers = if_else(rowname %in% c(14, 57, 58, 249, 280, 16, 19, 20, 118), 1, 0)) %>% 
  filter(lrti_incidence < 0.7) %>% 
  filter(filter_outliers == 0)

@@

# amb data summary
# amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti
ambient_summary <- analysis_data %>% 
  filter(area == "NN") %>% 
  mutate(new_var = amb_pm_lrti) %>% 
  group_by(treatment) %>% 
  filter(!is.na(new_var)) %>% 
  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% 
  summarize(mean_amb = mean(new_var, na.rm = TRUE),
            sd_amb = sd(new_var, na.rm = TRUE),
            n_amb = n(),
            min_amb = min(new_var, na.rm = TRUE),
            median_amb = median(new_var, na.rm = TRUE),
            max_amb = max(new_var, na.rm = TRUE))
ambient_summary
```

```{r}
# Histogram for distribution of data
events_plot <- analysis_data %>% 
  ggplot() +
    geom_histogram(aes(lrti_events)) +
    theme_light() +
    facet_wrap(~treatment)
events_plot

inc_plot <- analysis_data %>% 
  ggplot() +
    geom_histogram(aes(lrti_incidence)) +
    theme_light() +
    facet_wrap(~treatment)
inc_plot

events_di_plot <- analysis_data %>% 
  ggplot() +
    geom_histogram(aes(lrti_events_di)) +
    facet_wrap(~treatment)
events_di_plot
```

```{r}
# Summarize initial models, check overdispersion and other diagnostic criteria
library(MASS)

# Simple model with offset
nb_results <- glm.nb(lrti_events ~ treatment + offset(log_person_time_at_risk), 
                     data = analysis_data)

summary(nb_results)
anova(nb_results, test="Chisq")

@@

summary(poisson_results)
anova(poisson_results, test="Chisq")

@@

# Other diagnostic plots and tests
# https://stats.stackexchange.com/questions/70558/diagnostic-plots-for-count-regression

library(vcd)
fit <- goodfit(analysis_data$lrti_events)
summary(fit) 
rootogram(fit)
Ord_plot(analysis_data$lrti_events)
distplot(analysis_data$lrti_events, type = "nbinom")

library(AER)
dispersiontest(poisson_results, trafo=1) # test indicates poisson model is overdispersed

library(pscl)
odTest(nb_results) # test indicates poisson model is overdispersed
# According to these results, data is overdispersed
# Variance is much higher than the mean for LRTI incidence
# Assumption of Poisson is variance = mean, so recommend using negative binomial model

mod2 <- zeroinfl(lrti_events ~ treatment + offset(log_person_time_at_risk), 
                 data = analysis_data, dist="poisson")
AIC(poisson_results, mod2)
# Above results indicate zero-inflated poisson model is a better fit than poisson model

mod2 <- zeroinfl(lrti_events ~ treatment + offset(log_person_time_at_risk), 
                 data = analysis_data, dist="negbin")
AIC(nb_results, mod2)
# Above results indicate neg bin model is a better fit than zero-inflated neg bin model
# Since data is also overdispersed (neg bin model better fit than poisson),
# a neg bin model seems to be the best fit overall

# plot residuals
res <- residuals(nb_results, type="deviance")
plot(predict(nb_results), res)
abline(h=0, lty=2)
qqnorm(res)
qqline(res)

library(faraway)
halfnorm(residuals(nb_results))

library(boot)
glm.diag.plots(nb_results)

# Check for influential datapoints
library(car)
nb_influence <- influencePlot(nb_results)
influencePlot(poisson_results)

@@

influence(nb_results)$coef

qqnorm(residuals(nb_results))
qqnorm(residuals(poisson_results))
```

@@

@@

```{r}
# Run mixed model and diagnostic plots from above

negbin_results <- glmer.nb(lrti_events ~ treatment_assigned + child_age + (1 | home:cohort:area) + 
                           offset(log_person_time_at_risk), data = analysis_data)

summary(negbin_results)
tidy(negbin_results, conf.int = TRUE)

# plot residuals
res <- residuals(negbin_results, type="deviance")
plot(predict(negbin_results), res)
abline(h=0, lty=2)
qqnorm(res)
qqline(res)

plot(residuals(negbin_results) ~ fitted(negbin_results), xlab="Fitted", ylab="Residuals")
abline(h=0)

library(faraway)
halfnorm(residuals(negbin_results))

```
